Mutual Information Plots

### 2025-10-17 mutual_information_curves.qmd
Zach's slides

```{r}
library(gtools)
library(tidyverse)
partial_kendall <- function(vec, k, weights = NULL){
  n <- length(vec) 
  if (k == n){k <- n-1} 
  ref <- c(1:n)

@@

perms_list <- permn(1:5)

# Apply weighted_kendall to each permutation
results <- sapply(perms_list, function(p) {
  weighted_kendall(rhat = p, k = 5, wtype = "symmetric_salience")
})

# Put into a dataframe
kendall_df <- data.frame(
  permutation = sapply(perms_list, paste, collapse = ""),
  weighted_kendall = results
)

kendall_df

sum(kendall_df$weighted_kendall)
mean(kendall_df$weighted_kendall)
```

>>>>

```{r}
library(gtools)
library(tidyverse)
library(combinat)

squared_error <- function(vec, weights = NULL) {
  n <- length(vec)
  ref <- 1:n

  # Default weights = 1 for each element
  if (is.null(weights)) {
    weights <- rep(1, n)
  }

  # Initialize sum
  summ <- 0

  # Accumulate weighted squared error
  for (i in 1:n) {
    summ <- summ + (weights[i] * (vec[i] - ref[i]))^2
  }

  # Handle near-zero sum safely
  if (abs(summ) < 1e-10) {
    out <- 1
  } else {
    out <- 1 / summ
  }

  return(out)
}

partial_kendall <- function(vec, k, weights = NULL){
  n <- length(vec) 
  if (k == n){k <- n-1} 
  ref <- c(1:n)

@@

make_weights <- function(k) {
  w <- rep(0, 4)
  w[1:k] <- 1
  return(w)
}

# Store all results
all_results <- list()

# Loop over k = 1 to 4
for (k in 1:4) {
  weights <- make_weights(k)

  # Weighted Kendall (using provided function)
  kendall_results <- sapply(perms_list, function(p) {
    weighted_kendall(rhat = p, k = 4, wtype = "symmetric_salience")
  })

  # Squared error using current weights
  sq_results <- sapply(perms_list, function(p) {
    squared_error(vec = p, weights = weights)
  })

  # Combine into one dataframe for this k
  rhat_df <- data.frame(
    k = k,
    permutation = sapply(perms_list, paste, collapse = ""),
    weighted_kendall = kendall_results,
    squared_error = sq_results
  )

  all_results[[k]] <- rhat_df
}

# Combine across all k
rhat_df <- bind_rows(all_results)
```
